---
title: "Course Cleaning"
output: html_notebook
---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  error = FALSE, 
  cache = FALSE
)
```

Read in data.

```{r}
library(dplyr)
library(ggplot2)
library(stringr)

theme_set(theme_classic())


courses_df <- readr::read_csv(
  file.path('intermediate', 'courses.csv'), 
  show_col_types = FALSE
)
```

Define a sampling function to quickly peak at the data:

```{r}
peak <- function(df) {
  return(slice_sample(df, n = 100) %>% select(course))
}
```


There are 788 unique degrees we found and put in the Google spreadsheet. Of the 
788, we coded course information for 672 of them. Most of the course codes 
include multiple courses on multiple lines. Our starting point is to separate 
each line as a potential course for a degree.

```{r}
nrow(courses_df) 
```

There are 36,727 potential degrees. A lot of these are blank lines or 
miscellaneous formatting. We will attempt to clean and remove as many
as we can. 

```{r}
courses_df <-
  courses_df %>%
  filter(!is.na(course))

nrow(courses_df)
```

19,454 "courses" were just blank lines. Now we have 17,273 courses. 

Some of the courses include course descriptions. Some of the courses are only
descriptions, because the course name was on a previous line. We probably want
to get rid of the longest course texts. 

Here are courses longer than 80 characters:

```{r}
courses_df %>%
  filter(str_length(course) > 80) %>%
  select(course)
```

I think it is valid to drop anything longer than 80 characters. 

```{r}
courses_df <-
  courses_df %>%
  filter(str_length(course) < 80)

nrow(courses_df)
```

1,679 courses were course descriptions. Removing them leaves us with 15,594 courses to clean.

Let's take a random sample of course text to pick our next cleaning step.

```{r}
peak(courses_df)
```

I want to remove any parentheses or brackets and whatever is in them.

```{r}
pattern <- regex('[\\(\\[].*[\\)\\]]')

courses_df <-
  courses_df %>%
  mutate(course = str_remove(course, pattern))
```

Take a sample to see how that worked.

```{r}
peak(courses_df)
```

Let's take a look at the very shortest course names.

```{r}
courses_df %>%
  filter(str_length(course) < 10) %>%
  arrange(str_length(course)) %>%
  select(course)
```
I saw nothing under length 7 that was worth saving.

```{r}
courses_df <-
  courses_df %>%
  filter(str_length(course) >= 7)

nrow(courses_df)
```

There were 1,159 courses under length 7 that we removed, bringing our total
down to 14,435.

Sampling to review:

```{r}
peak(courses_df)
```

I want to get rid of the course code and department abbreviation at the 
beginning of course text. 

Previewing what would removed if I targetted words with mixed letters and 
numbers at the beginning of course text:

```{r}
pattern <- regex('^\\w*\\d[\\w\\d[:punct:]]+[[:space:]]?[[:punct:]]?[[:space:]]?')

courses_df %>% 
  mutate(extracted = str_extract(course, pattern)) %>%
  filter(!is.na(extracted)) %>%
  select(extracted)
```

That looks good to me. Let's remove it.

```{r}
courses_df <-
  courses_df %>%
  mutate(course = str_remove(course, pattern))
```

And sample the course after this cleaning:

```{r}
peak(courses_df)
```

Going to try again because there are still course codes etc. left starting 
the course stings.

Here's a preview:

```{r}
pattern <- regex('^[:alpha:]{2,5}[:blank:]?[:digit:]{3,4}', ignore_case = TRUE)

courses_df %>% 
  mutate(extracted = str_extract(course, pattern)) %>%
  filter(!is.na(extracted)) %>%
  select(extracted)
```

Let's do it.

```{r}
courses_df <- 
  courses_df %>%
  mutate(course = str_remove(course, pattern))
```

And sample the results:

```{r}
peak(courses_df)
```

Looking good! 

Let's get rid of any punctuation characters at the start of 
courses. First I'm going to trim any leading blank spaces, then trim the punctuation, then trim any remaining whitespace.

OK, now to remove the leading punctuation.

```{r}
pattern <- regex('^[[:punct:]]')

courses_df <-
  courses_df %>%
  mutate(course = str_trim(course, side = 'left'),
         course = str_remove(course, pattern),
         course = str_trim(course, side = 'left'))
```

Sampling the results:

```{r}
peak(courses_df)
```
Looking pretty good!

Bet we have a bunch of short stubs if we sort by length.

```{r}
courses_df %>%
  filter(str_length(course) < 7) %>%
  select(course) %>%
  arrange(str_length(course)) 
```

Yeah, it's a bunch of blanks.

Let's remove them and see what we have left.

```{r}
courses_df <-
  courses_df %>%
  filter(str_length(course) >= 7)

nrow(courses_df)
```

There were 824 short courses after our recent cleaning steps. We're now down 
to 13,611 courses.

Sampling the results:

```{r}
peak(courses_df)
```

Let's try removing some stop words.

I'm thinking:
- required
- elective
- course
- courses
- credits
- credit
- choose
- following
- summer
- fall
- winter
- spring

Some of these indicate a phrase which is not a course at all, like "Required courses".

Say the course text was "Required Statistics Courses". If we pull out "required" and
"courses", we're left with "statistics" which sounds like the name of a course. But it's not.

In that case, we just want to delete the whole course text if we find "required". 

This includes:
- course
- courses
- required
- requirements
- elective
- choose
- following

For other words, we just want to remove the word but leave the rest of the text. 

Example: "Data Mining 3 Credits". We want to remove "Credits" (and later, "3") but 
we should keep "Data Mining". 

These words are:
- credits
- credit
- cr
- hour
- hr
- hours

I'm not sure what to do with fall/winter/spring/summer yet. 

Starting by deleting full course text:

```{r}
pattern <- regex('course|courses|required|requirements|elective|choose|following', ignore_case = T)

courses_df %>%
  filter(str_detect(course, pattern)) %>%
  select(course)
```

This looks great. Let's get rid of 'em.

```{r}
courses_df <- 
  courses_df %>%
  filter(!str_detect(course, pattern))

nrow(courses_df)
```

This removed 1,427 courses that were not actually courses, more likely 
headers and stuff from course sections of websites. We're down to 12,184.

How's it look?

```{r}
peak(courses_df)
```

Now let's remove those other words but not delete the whole course text.

```{r}
pattern <- regex('credits|credit|hour|hours|semester|units', ignore_case = T)

courses_df %>%
  mutate(course = str_remove_all(course, pattern)) %>%
  select(course)
```

That looks good. Let's do it.

```{r}
courses_df <-
  courses_df %>%
  mutate(course = str_remove_all(course, pattern)) 

nrow(courses_df)
```

So we pulled some stop words, but unlike the previous step we did not remove 
entire rows containing the stop words. I figure there are some newly short 
courses that we can remove as a result. 

```{r}
courses_df <-
  courses_df %>%
  filter(str_length(course) >= 7)

nrow(courses_df)
```

OK, 194 short courses removed. Down to 11,990 courses. 

Let's take a random slice and see how it's looking:

```{r}
peak(courses_df)
```

Pretty clean! One of the last things to do is get rid of numbers.

Preview of what we will extract:

```{r}
pattern <- regex('\\w*\\d\\w*')

courses_df %>%
  mutate(extracted = str_extract(course, pattern)) %>%
  filter(!is.na(extracted)) %>%
  select(extracted, course) 
```

That's good. Let's do it.

```{r}
courses_df <-
  courses_df %>%
  mutate(course = str_remove(course, pattern))
```

```{r}
peak(courses_df)
```

OK so I think I went as far as I could go in R. 

Exporting the data so that I can manually clean it up row by row in a spreadsheet.

```{r}
readr::write_csv(courses_df, file = file.path('intermediate', 'courses_pre_manual_clean.csv'))
```

Continued in `data-raw/course_cleaning2.Rmd`!
