---
title: "course_cleaning2"
output: html_document
---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  error = FALSE, 
  cache = FALSE
)
```

In the first course cleaning step we made the 
`intermediate/courses_pre_manual_clean.csv` file.

I opened it up in [Google sheets](https://docs.google.com/spreadsheets/d/178EsntNONoUQm2Y49HyBacRXxSWlGXOtbF-alaODVvY/edit#gid=1901186123) and inspected it row by row. 

The main things I removed:
- department/course codes ("BUSADM345")
- credit numbers ("3 cr")
- non-course text ("Data Science Concentration:")
- 2 courses in one row ("Data science OR Statistics")
  + I removed the second course in this case
- Misc. punctuation left over from cleaning
- Instructor names

```{r}
library(dplyr)
library(ggplot2)
library(stringr)

theme_set(theme_classic())


courses_df <- readr::read_csv(
  file.path('intermediate', 'courses_pre_manual_clean.csv'), 
  show_col_types = FALSE
)

courses_df2 <- readr::read_csv(
  file.path('intermediate', 'courses_post_manual_clean.csv')
)
```

Removing blanks resulting from the cleaning:

```{r}
courses_df2 <-
  courses_df2 %>%
  select(degree_id, course) %>%
  filter(!is.na(course))
```

So before cleaning we had this many courses:

```{r}
nrow(courses_df)
```

And now we have:

```{r}
nrow(courses_df2)
```

So we dropped 1,733 rows of non-course text. 

There's still a little bit of cleaning to do. Mostly just trim leading and 
trailing blank spaces around the course text. 

```{r}
courses_df2 <- mutate(courses_df2, course = str_trim(course, side = 'both'))
```

```{r}
slice_sample(courses_df2, n = 100) %>% select(course)
```

Still have a few issues:

- some text is all caps
- have a lot of "I" and "II" ("Physics I or II")
- there are course duplicates per degree

Lowering all course text:

```{r}
courses_df2 <-mutate(courses_df2, course = str_to_lower(course))
```

```{r}
slice_sample(courses_df2, n = 100) %>% select(course)
```

Let's try to get rid of some of the roman numerals.

```{r}
courses_df2 <- courses_df2 %>%
  mutate(course = str_remove(course, pattern = regex(' i+$')))
```

```{r}
slice_sample(courses_df2, n = 100) %>% select(course)
```

OK, now to remove duplicate courses per degree:

```{r}
courses_df2 <- distinct(courses_df2, degree_id, course)
```

This leaves us with:

```{r}
nrow(courses_df2)
```

9,184 courses! That's pretty manageable. 

Let's save it and call it a day.

```{r}
readr::write_csv(courses_df2, file = file.path('..', 'data', 'courses.csv'))
```

